---
title: "General Social Survey"
author: "Nicolas Rizo"
date: "October 7 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data. and chi-squared test for inference

$H_0$ : There is no relationship between gun ownership and whether one is in favor of capitol punishment

$H_A$ : There is a relationship between gun ownership and whether one is in favor of capitol punishment

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+OWNGUN, data=mydata4, main= 'Punishment and Gun Ownership', type='percent', xlab='Captial Punishment')
```
This barchart shows that there is a higher percentage of people who oppose Capitol Punishment and do not own a  gun than those who are in favor. 


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1= xtabs(~CAPPUN+OWNGUN, data=mydata4)
rowPerc(tab1)
```
This table shows that more people do not own a gun whether or not they are in favor of Capitol Punishment 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```

The p-value for the chisq test is 0.02 < ..05 so we reject the null hypothesis and conclude we have evidence to support the contention that gun ownership and whether one favors capital punishment are related

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

For this analysis we will use bar graphs and tables to summarize the data. and chi-squared test for inference

$H_0$ : There is no relationship between belief in the Big Bang Theory and whether one believes in afterlife

$H_A$ :There is a relationship between belief in the Big Bang Theory and whether one believes in afterlife
 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main= 'Bigbang and Postlife', type = 'percent')
```
This barchart shows that the percentage of those who believe in postlife as well as the big bang theory is about the same. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1= xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(tab1)
```
This numerical table shows that there is a higher number of those who believe in postlife with roughly 82 also believing in the Bigbang and about 92 who do not believe in the BigBang

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```
The p-value for the chisq test is 0.0801 > ..05 so we fail to reject the null hypothesis and conclude we do not have evidence to support the contention that the belief in the BigBang theory and Aferlife are related.

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

For this analysis we will use bar graphs and tables to summarize the data. and chi-squared test for inference

$H_0$ : There is no relationship between one's opinion on the death penalty and their race.

$H_A$ :There is a relationship between one's opinion on the death penalty and their race.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main= 'Punishment and Race', type='percent', xlab='Captial Punishment')
```
This barchart shows that their is a higher percentage of Black's who are in opposition than those in favor. For White's the percentage of those in favor and in  oppisition is roughly the same

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1= xtabs(~CAPPUN+RACE, data=mydata4)
rowPerc(tab1)
```
This numerical table shows that there are more whites in favor with 79 than in opposition with about 68. There is also a higher number of those in opposition for the blacks, with 27 than those in favor, with 8.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```
The p-value for the chisq test is  9e-04 > ..05 so we reject reject the null hypothesis and conclude that their is a relation ship between one's race and their opinion on capitol punishment


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Is there a relationship between gun ownership and whether one is in favor of capitol punishment?
There was a P-Value of 0.02 < ..05 so we reject the null hypothesis and conclude we have evidence to support the contention that gun ownership and whether one favors capital punishment are related

Is there a relationship between belief in the Big Bang Theory and whether one believes in afterlife? The P-value was 0.0801 > ..05 so we fail to reject the null hypothesis and conclude we do not have evidence to support the contention that the belief in the BigBang theory and Aferlife are related.

Is there a relationship between one's opinion on the death penalty and their race?
The P-value was 9e-04 > ..05 so we reject reject the null hypothesis and conclude that their is a relation ship between one's race and their opinion on capitol punishment
